---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cm

<!-- badges: start -->
<!-- badges: end -->

Proof of concept (poc) for a contact matrix class that could underpin various
packages (e.g. [socialmixr](https://cran.r-project.org/package=socialmixr)).

## Installation

You can install cm from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("epi-helpers/cm")
```

## Overview

### A contact matrix class

The underlying `contact_matrix` class is very simple with the only requirements
on the underlying data being that it must be a numeric matrix of equal
dimensions (i.e. an equal number of rows and columns). Encoded with this data is
an additional `from` attribute which conveys to a user whether the matrix should
be read as from "rows to columns" or from "columns to rows".

### Subclasses 
`cm` also provides two contact matrix subclasses based on those discussed in
\cite{Klepec et al}:

* a `social_contact_matrix`; and
* a `population_contact_matrix`.

The motivation for explicit classes is to avoid confusion on behalf of package
users/developers (e.g. are contacts represented as from "rows to columns"
or "columns to rows"; is this matrix symmetric with respect to population
weightings?).

The subclasses deliberately simple. For social contact matrices these are just a
subclass of a contact matrix. For population contact matrices they require an
additional attribute that encapsulates the population weightings
that give symmetry, that is \eqn{C_{ij} * w_{j} == C_{ji} * w_{i}}, where
\eqn{w_{k}} represents the population in group *k*.

### Example

We use data from [socialmixr](https://cran.r-project.org/package=socialmixr)
to understand how we could utilise explicit matrix classes. Currently
construction is slightly long-winded as I still need to provide general coercion
functions for ready formed matrices (remember this is just a poc).

```{r}
library(cm)
library(socialmixr)

# first obtain data for the social contact matrix and underlying population
dat <- contact_matrix(
    polymod,
    countries = "United Kingdom",
    age.limits = c(0, 1, 5, 15)
)
dat <- dat$matrix
dat2 <- contact_matrix(
    polymod,
    countries = "United Kingdom",
    age.limits = c(0, 1, 5, 15),
    symmetric = TRUE
)
popn <- dat2$demography$population

# currently we require long format data
(dat <- as.data.frame(as.table(dat)))

# create a social contact matrix
(social_mat <- with(
    dat,
    social_contact_matrix(age.group, contact.age.group, Freq, from = "rows")
))

# convert to a population contact matrix
(population_mat <- as_population_matrix(social_mat, population = popn))

# compare this to the data already loaded from socialmixr
(original <- dat2$matrix)
dimnames(original) <- NULL
all.equal(as.matrix(population_mat), original)
```
